Type Refinements for the Working Class

### 2015-11-20 paper.tex
Move the formal definitions of higher order judgments into appendix (#5)

For any abt signature $\Match{\Sigma}{\Pair{\Sorts}{\Operators}}$, we can deploy
a generic logical framework. Recall from~\cite{sterling-morrison:2015} that
$\SCtx$ is the category of symbol-contexts and their injective, sort-preserving
renamings. Then, a \emph{judgment} shall be interpreted as an object in the presheaf
category $\Define{\mathbb{J}}{\Sets^\SCtx}$; in other words, a
judgment is identified with the intensional set of its renaming-sensitive
derivations. $\mathbb{J}$, like all presheaf categories, is a topos.

\begin{definition}[Hypothetical Judgment]
  For two judgments $\mathcal{J}_1$ and $\mathcal{J}_2$, we can form the
  hypothetical judgment $\HypJ{\mathcal{J}_2}{\mathcal{J}_1}$ using the
  exponential construction~\cite[p.~46]{maclane-moerdijk:1992}, which expresses
  the semantic consequence of $\mathcal{J}_2$ from $\mathcal{J}_1$:
  \begin{align*}
    \ADefine{
      \left(\HypJ{\mathcal{J}_2}{\mathcal{J}_1}\right)(\Upsilon)
    }{
      \left({\mathcal{J}_2}^{\mathcal{J}_1}\right)(\Upsilon)
    }\\
    &\cong
    \OMode{
      \Hom{\mathbb{J}}{
        \Yoneda{\Upsilon}\times\mathcal{J}_1
      }{
        \mathcal{J}_2
      }
    }\\
    &\cong
    \OMode{
      \int_{\Upsilon'}{{\Upsilon\inj\Upsilon'}
        \Rightarrow{{{\mathcal{J}}_{2}(\Upsilon')}^{{\mathcal{J}}_{1}(\Upsilon')}}}
    }
  \end{align*}

  In other words, the hypothetical judgment
  $\HypJ{\mathcal{J}_2}{\mathcal{J}_1}$ is evident when there shall forevermore
  be an effective transformation of evidence for the antecedent into evidence
  for the consequent. It is important to keep in mind that this is a
  \emph{semantic} consequence and therefore expresses admissibilities, in
  contrast to the \emph{logical} consequence which expresses
  derivabilities~\cite{harper:2016}. We will not have any need for the latter.

@@

  For any diagram $\Of{\mathcal{J}}{\mathbb{J}^{\square\Theta}}$, we can define
  the general judgment $\GenJ{\Theta}{\mathcal{J}}$ as follows, where
  $\Of{\Yoneda{-}}{\OpCat{\SCtx}\ra\mathbb{J}}$ is the Yoneda Embedding:
  \begin{align*}
    \ADefine{
      \left(\GenJ{\Theta}{\mathcal{J}}\right)(\Upsilon)
    }{
      \Hom{\mathbb{J}}{
        \Yoneda{\Upsilon}
      }{
        \varprojlim\mathcal{J}
      }
    }\\
    &\cong
    \OMode{
      \int_{(\Upsilon',\,\vec{E})}{\Upsilon\inj\Upsilon'\Rightarrow{\mathcal{J}(\vec{E})(\Upsilon')}}
    }
  \end{align*}

   Intuitively, the general judgment $\GenJ{\Theta}{\mathcal{J}}$ is evident
   when the family of judgments $\mathcal{J}$ shall forevermore have a global
   section (i.e.\ for any substitution $\Member{\vec{E}}{\square\Theta}$,
   $\mathcal{J}(\vec{E})$ shall be evident). Recall that metavariables have
   \emph{valences} rather than sorts; as such, the general judgment allows us
   to quantify over terms of higher type.

@@

\begin{definition}[Internal Judgments]

  It often happens that the evidence for a judgment is the knowledge of the
  meaning of a new form of judgment. This can be neatly expressed in our topos
  using the subobject classifier $\Omega$, where the evidence of $\Omega$ at
  $\Upsilon$ gives rise to an \emph{internal judgment} which is defined at all
  $\Upsilon'$ for which we have $\Of{\rho}{\Upsilon\inj\Upsilon'}$.

  When asserting the existence of an internal judgment, in most cases we will
  say something like ``To know $\mathcal{J}$ is to have a judgment
  $\mathcal{J'}$'' rather than appealing directly to the subobject classifier.
\end{definition}

@@

 \begin{remark}
   Formally, the evaluation relation is given in our logical framework by a
   global internal judgment $\Of{\Downarrow_\tau}{\int_\SCtx
   \Omega^{\Exprs_\tau\times\Values_\tau\times\NaturalsPsh}}$, where
   $\NaturalsPsh$ is the constant presheaf $\MapsTo{\Upsilon}{\Naturals}$.
 \end{remark}

\ifdraft{}{
  \nocite{maclane:1971}
  \bibliographystyle{abbrv}
  \bibliography{refs}
}

>>>>

For any abt signature $\Match{\Sigma}{\Pair{\Sorts}{\Operators}}$, we can
deploy a generic logical framework equipped with higher-order judgments. A
judgment is always defined using a \emph{meaning explanation}, which is a
specification of the conditions under which it may be asserted (sc.\ its
evidence) \cite{martin-lof:1996}.

\paragraph{Renaming convention} Because the evidence of a judgment will often
contain symbols $u,v,\dots$ we informally adopt the convention that the
evidence of a judgment shall always be subject to fresh renamings of its free
symbols (i.e.\ the validity of a judgment shall not depend on choice of names).
This apparatus is developed formally in Appendix~\ref{appendix:lf}, where the
logical framework is cast as a topos of covariant presheaves on symbol contexts
$\SCtx$.

\begin{definition}[Hypothetical Judgment]
  For two judgments $\mathcal{J}_1$ and $\mathcal{J}_2$, we can form the
  hypothetical judgment $\HypJ{\mathcal{J}_2}{\mathcal{J}_1}$, whose evidence
  shall be an effective transformation of any evidence for $\mathcal{J}_1$ into
  evidence for $\mathcal{J}_2$. It is important to keep in mind that this is a
  \emph{semantic} consequence and therefore expresses admissibilities, in
  contrast to the \emph{logical} consequence which expresses
  derivabilities~\cite{harper:2016}. We will not have any need for the latter.

@@

  For any $\square{\Theta}$-indexed family of judgments $\mathcal{J}$, we can
  express the general judgment $\GenJ{\Theta}{\mathcal{J}}$ as being evident
  when $\mathcal{J}(\vec{E})$ is evident for each
  $\Of{\vec{E}}{\square\Theta}$; in other words, the evidence of the general
  judgment is an effective transformation of $\Theta$-substitutions to their
  fibres in $\mathcal{J}$.  Recall that metavariables have \emph{valences}
  rather than sorts; as such, the general judgment allows us to quantify over
  terms of higher type.

@@

@@

\newpage
\begin{appendices}
\section{Formal Definition of the Logical Framework}
\label{appendix:lf}

As noted earlier, we must be cautious in the definitions of both atomic and
higher-order judgments that their evidence shall always respect the
\emph{renaming convention}. We do so explicitly in this appendix by developing
our logical framework as a presheaf topos over symbol contexts.

A judgment is, then, an object in the presheaf category
$\Define{\mathbb{J}}{\Sets^\SCtx}$; in other words, judgments are identified
with the intensional set of its renameable evidence. $\mathbb{J}$, like all
presheaf categories, is a topos. The task at hand, then, is to show how to
replace the informal meaning explanations that we gave in
section~\ref{section:ambient-lf} with precise definitions in the topos
$\mathbb{J}$.

\subsection{Hypothetical Judgment}

Hypothetical judgment $\HypJ{\mathcal{J}_2}{\mathcal{J}_1}$ is defined using the
exponential construction~\cite[p.~46]{maclane-moerdijk:1992}, which expresses
the semantic consequence of $\mathcal{J}_2$ from $\mathcal{J}_1$:
\begin{align*}
  \ADefine{
    \left(\HypJ{\mathcal{J}_2}{\mathcal{J}_1}\right)(\Upsilon)
  }{
    \left({\mathcal{J}_2}^{\mathcal{J}_1}\right)(\Upsilon)
  }\\
  &\cong
  \OMode{
    \Hom{\mathbb{J}}{
      \Yoneda{\Upsilon}\times\mathcal{J}_1
    }{
      \mathcal{J}_2
    }
  }\\
  &\cong
  \OMode{
    \int_{\Upsilon'}{{\Upsilon\inj\Upsilon'}
      \Rightarrow{{{\mathcal{J}}_{2}(\Upsilon')}^{{\mathcal{J}}_{1}(\Upsilon')}}}
  }
\end{align*}

In other words, the hypothetical judgment $\HypJ{\mathcal{J}_2}{\mathcal{J}_1}$
is evident when there shall ``forevermore'' (i.e. for any applicable renaming
of symbols) be an effective transformation of evidence for the antecedent into
evidence for the consequent.

\subsection{General Judgment}

For any diagram $\Of{\mathcal{J}}{\mathbb{J}^{\square\Theta}}$, we can define
the general judgment $\GenJ{\Theta}{\mathcal{J}}$ as follows:
\begin{align*}
  \ADefine{
    \left(\GenJ{\Theta}{\mathcal{J}}\right)(\Upsilon)
  }{
    \Hom{\mathbb{J}}{
      \Yoneda{\Upsilon}
    }{
      \varprojlim\mathcal{J}
    }
  }\\
  &\cong
  \OMode{
    \int_{(\Upsilon',\,\vec{E})}{\Upsilon\inj\Upsilon'\Rightarrow{\mathcal{J}(\vec{E})(\Upsilon')}}
  }
\end{align*}

 Intuitively, the general judgment $\GenJ{\Theta}{\mathcal{J}}$ is evident
 when the family of judgments $\mathcal{J}$ shall forevermore have a global
 section (i.e.\ for any substitution $\Member{\vec{E}}{\square\Theta}$,
 $\mathcal{J}(\vec{E})$ shall be evident).

\subsection{Internal Judgments}
It often happens that the evidence for a judgment is the knowledge of the
meaning of a new form of judgment; this is the case when we say ``To know
$\mathcal{J}$ is to know \emph{the meaning of} another judgment
$\mathcal{J}'$.'' This can be neatly expressed in our topos using the subobject
classifier $\Omega$, where the evidence of $\Omega$ at $\Upsilon$ gives rise to
an \emph{internal judgment} which is defined at all $\Upsilon'$ for which we
have $\Of{\rho}{\Upsilon\inj\Upsilon'}$.

\end{appendices}

\ifdraft{}{
  \newpage
  \nocite{maclane:1971}
  \bibliographystyle{abbrv}
  \bibliography{refs}
}
